---
title: "main_rf"
author: "Luyao Sun"
date: "11/2/2020"
output: html_document
---

---
title: "Main"
author: "Chengliang Tang, Yujie Wang, Diane Lu, Tian Zheng"
output:
  pdf_document: default
  html_notebook: default
---

In your final repo, there should be an R markdown file that organizes **all computational steps** for evaluating your proposed Facial Expression Recognition framework. 


This file is currently a template for running evaluation experiments. You should update it according to your codes but following precisely the same structure. 

```{r message=FALSE}
if(!require("EBImage")){
  install.packages("BiocManager")
  BiocManager::install("EBImage")
}
if(!require("R.matlab")){
  install.packages("R.matlab")
}
if(!require("readxl")){
  install.packages("readxl")
}

if(!require("dplyr")){
  install.packages("dplyr")
}
if(!require("readxl")){
  install.packages("readxl")
}

if(!require("ggplot2")){
  install.packages("ggplot2")
}

if(!require("caret")){
  install.packages("caret")
}

if(!require("glmnet")){
  install.packages("glmnet")
}

if(!require("WeightedROC")){
  install.packages("WeightedROC")
}

library(R.matlab)
library(readxl)
library(dplyr)
library(EBImage)
library(ggplot2)
library(caret)
library(glmnet)
library(WeightedROC)
library(randomForest)
library(tidyverse)
library(tidymodels)
library(doParallel)
```

### Step 0 set work directories
```{r wkdir, eval=FALSE}
set.seed(2020)
#setwd("~/Project3-FacialEmotionRecognition/doc")
# here replace it with your own path or manually set it in RStudio to where this rmd file is located. 
# use relative path for reproducibility
```

Provide directories for training images. Training images and Training fiducial points will be in different subfolders. 
```{r}
train_dir <- "../data/train_set/" # This will be modified for different data sets.
train_image_dir <- paste(train_dir, "images/", sep="")
train_pt_dir <- paste(train_dir,  "points/", sep="")
train_label_path <- paste(train_dir, "label.csv", sep="") 
```

### Step 1: set up controls for evaluation experiments.

In this chunk, we have a set of controls for the evaluation experiments. 

+ (T/F) cross-validation on the training set
+ (T/F) reweighting the samples for training set 
+ (number) K, the number of CV folds
+ (T/F) process features for training set
+ (T/F) run evaluation on an independent test set
+ (T/F) process features for test set

```{r exp_setup}
# run.cv <- TRUE # run cross-validation on the training set
# sample.reweight <- TRUE # run sample reweighting in model training
# K <- 5  # number of CV folds
 run.feature.train <- TRUE # process features for training set
 run.test <- TRUE # run evaluation on an independent test set
 run.feature.test <- TRUE # process features for test set
```

Using cross-validation or independent test set evaluation, we compare the performance of models with different specifications. In this Starter Code, we tune parameter lambda (the amount of shrinkage) for logistic regression with LASSO penalty.

```{r model_setup}
# lmbd = c(1e-3, 5e-3, 1e-2, 5e-2, 1e-1)
# model_labels = paste("LASSO Penalty with lambda =", lmbd)
```

### Step 2: import data and train-test split 
```{r}
#train-test split
info <- read.csv(train_label_path)
n <- nrow(info)
n_train <- round(n*(4/5), 0)
train_idx <- sample(info$Index, n_train, replace = F)
test_idx <- setdiff(info$Index, train_idx)
```

If you choose to extract features from images, such as using Gabor filter, R memory will exhaust all images are read together. The solution is to repeat reading a smaller batch(e.g 100) and process them. 
```{r}
n_files <- length(list.files(train_image_dir))

# image_list <- list()
# for(i in 1:100){
#    image_list[[i]] <- readImage(paste0(train_image_dir, sprintf("%04d", i), ".jpg"))
# }
```

Fiducial points are stored in matlab format. In this step, we read them and store them in a list.
```{r read fiducial points}
#function to read fiducial points
#input: index
#output: matrix of fiducial points corresponding to the index
readMat.matrix <- function(index){
     return(round(readMat(paste0(train_pt_dir, sprintf("%04d", index), ".mat"))[[1]],0))
}

#load fiducial points
fiducial_pt_list <- lapply(1:n_files, readMat.matrix)
save(fiducial_pt_list, file="../output/fiducial_pt_list.RData")
```

### Step 3: construct features and responses
`feature.R` should be the wrapper for all your feature engineering functions and options. The function `feature( )` should have options that correspond to different scenarios for your project and produces an R object that contains features and responses that are required by all the models you are going to evaluate later. 
  
  + `feature.R`
  + Input: list of images or fiducial point
  + Output: an RData file that contains extracted features and corresponding responses

```{r feature}
source("../lib/feature.R")
tm_feature_train <- NA
if(run.feature.train){
  tm_feature_train <- system.time(dat_train <- feature(fiducial_pt_list, train_idx))
  save(dat_train, file="../output/feature_train.RData")
}else{
  load(file="../output/feature_train.RData")
}

tm_feature_test <- NA
if(run.feature.test){
  tm_feature_test <- system.time(dat_test <- feature(fiducial_pt_list, test_idx))
  save(dat_test, file="../output/feature_test.RData")
}else{
  load(file="../output/feature_test.RData")
}

# dim(dat_train)
# ##[1] 2400 6007
# dim(dat_test)
# ##[1]  600 6007
```

### Step 4: Train a classification model with training features and responses
Call the train model and test model from library. 

`train.R` and `test.R` should be wrappers for all your model training steps and your classification/prediction steps. 

+ `train.R`
  + Input: a data frame containing features and labels and a parameter list.
  + Output:a trained model
+ `test.R`
  + Input: the fitted classification model using training data and processed features from testing images 
  + Input: an R object that contains a trained classifier.
  + Output: training model specification

+ In this Starter Code, we use logistic regression with LASSO penalty to do classification. 

```{r loadlib}
# source("../lib/train.R") 
# source("../lib/test.R")
```

#### Build rf model

```{r}
tree_rec <- recipe(label~., data = dat_train)

tune_spec <- rand_forest(
  mtry = tune(),
  trees = 1000,
  min_n = tune()
) %>%
  set_mode("classification") %>%
  set_engine("ranger")


tune_wf <- workflow() %>%
  add_recipe(tree_rec) %>%
  add_model(tune_spec)

set.seed(0)
trees_folds <- vfold_cv(dat_train)

#doParallel::registerDoParallel()

tune_res <- tune_grid(
  tune_wf,
  resamples = trees_folds,
  grid =5
)
```